Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-28 background.tex
2013.01.23 (Alex) Introduction, background and conclusion sections reworked.

In this section, we provide some background information on the types of
peer-to-peer architectures that have been implemented and deployed. We also
describe and motivate more formally the topology mismatch problem tackled by
many P2P researchers over the past several years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Overlay Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Circa 2000, the first P2P file-sharing systems introduced the \emph{servent}
concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server
and client to denote the twofold role of a node participating in the P2P
network. Such server-less systems have proven to be able to achieve outstanding
aggregate resource capacities as more and more participants join the system
without requiring additional expenditure for infrastructure. While certain
undesirable features such as 
free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
the distribution of illegal content and other \emph{socio-technical}
issues \cite{hughes_socp2p_2008}, have arisen, peer-to-peer systems have
continued to gain popularity throughout the years primarily because they have successfully supported
popular applications (i.e. file-sharing) amongst vast numbers of participating
users.

For over a decade, excitement over the immense potential of P2P systems has
generated a flurry of research and development, resulting in a wide range of
popular protocols, networks, and applications. As P2P systems evolved, three
main architectures have emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

Designers of \emph{centralized} architectures (see
Figure~\ref{figure:p2p-archs:centralized}) were the first to recognize that
requests for resources (e.g. CPU cycles or popular content) need not be sent to
a dedicated server but instead could be handled by the many hosts that already
posess the resources. The downside of this approach is that it requires a
centralized searching facility which becomes a single point of failure, a
scalability bottleneck, and a target of malicious attacks (e.g.,
\emph{Denial-of-Service (DoS) attacks}).

The most successful incarnation of the centralized approach is the file-sharing
application \emph{Napster}\cite{napster}. Napster maintained a \emph{central
index server} based on file lists provided by participating peers. The central
index server was queried by users and it returned pointers to the actual
content. Thus, by centralizing search while distributing downloads,  Napster
achieved a highly functional design that, at its height, attracted 26 million
users sharing approximately 80 million songs\cite{jmm_naptopusage_2001}.
Napter's  centralized nature was what ultimately brought it down after Recording
Industry Association of America (RIAA) filed a lawsuit for copyright infringemnt
of the intellectual property of artists whose music was distributed by the
Napster application. RIAA was able to point to the centralized indexing
component of Napster as the entity resposible for illegal song-trading.

The centralized approach was quickly replaced by architectures that distribute
both search and resource provision capabilities (see
Figure~\ref{figure:p2p-archs:unstructured-full}). In these \emph{decentralized}
architectures, resource placement within the overlay topology is
random~\cite{YG-M2002}. For this reason they are referred to as
\emph{unstructured}. The most important properties of such systems are that they
support inherent heterogeneity of peers, are highly resilient to peer failures,
and incur low maintenance overhead at handling the dynamics of peer
participation \cite{stutzbach_churn_2006}. In the literature, they are also
known as \emph{broadcast-based} systems, because they use message flooding among
peers to propagate search queries. A well-known example of the fully
decentralized
unstructured approach is Gnutella, in its first realization (v.0.4)
\cite{gnutellav04}.

A key disadvantage of the unstructured approach is that message flooding burdens
the network since queries travel within the network randomly, visiting nodes
that do not have the resource of interest and thus wasting bandwidth.  To save
on bandwidth, unstructured networks typically limit how long a query travels
within the overlay (via a time-to-live (TTL) parameter). When a query is
received by a node, it decreases the TTL by one unit. If the TTL has not reached
zero, the node forwards the query to its neighbors.  Otherwise, it drops the
query. While this approach prevents queries from visiting the entire network, it
does not guarantee that the resource of interest, (e.g., a rare recording 
of a musical group from the 1930s) will be found.

To improve search scope and scalability and to lighten network load, 
unstructured systems evolved into ``hierarchical'' systems that dynamically
assign indexing functions to special peers, called \emph{ultra-peers}  or
\emph{super-peers} (see
Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). Peers with reliable
network connections and/or high compute power became ultra/super-peers
facilitating less powerful peers (called \emph{leaf-nodes}) to find resources of
interest. Well-known examples of these hierarchical unstructured systems include
Gnutella v.0.6/Limewire~\cite{gnutella}, KaZaA~\cite{kazaa} and
Skype~\cite{skype}. While the hierarchical approach relieves network pressure as
search queries are flooded over a much smaller subset of peers (the
ultra/super-peers), the disadvantages of reduced search scope and inability to
locate rare resources remain.

To fully address the disadvantages of decentralized unstructured approaches,
\emph{decentralized structured} schemes have been proposed (see
Figure~\ref{figure:p2p-archs:structured}). The key goal of these systems is to
provide a self-organizing infrastructure for large-scale P2P applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
through a deterministic mechanism. The DHT provides a guaranteed bound on the
number of overlay routing hops that have to be crossed to locate a resource, even
in the case when only a single copy exists in the system.  This requires
$O \left( log n \right)$ hops, compared to unstructured networks that require
$O \left( n \right)$ to reliably locate any object. Unfortunately this
efficiency does not come without cost. Structured overlays can only support
exact-match queries (i.e., queries that identify resources by name) as opposed
to content-based retrieval provided by unstructured overlays. This means
that unstructured overlays can support more versatile resource location queries
such as keyword and partial matching searches. Several popular file-sharing
peer-to-peer applications have been based on DHTs including
Overnet/eDonkey~\cite{overnet}, Kademlia/Kad
Network~\cite{maymounkov_kademlia_2002}, some flavors of
BitTorrent~\cite{c_bittorrent_2003}, and many more.  Moreover, DHTs have sparked
a flurry of research activity and development for a large range of applications:
from distributed search engines to event notification infrastructures to
cloud-based platforms. We cite several examples in Section~\ref{section:intro}.

\begin{figure}[ht]
\centering
\subfigure[Centralized (SETI@Home or Napster). Peers query the centralized
entity (dotted lines) and then establish the connection to the data source.] {
  \includegraphics[scale=0.25]{img/pdf/centralized.pdf}
  \label{figure:p2p-archs:centralized}
}\qquad\qquad
\subfigure[Fully unstructured (Gnutella v.0.4). Data is distributed randomly,
the topology is also random and all participants are equal.] {
  \includegraphics[scale=0.25]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructured-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack). Certain
peers (bigger circles) shoulder specific responsibilities (e.g., data
indexing).] {
  \includegraphics[scale=0.25]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured (Chord and Kademlia). Peers and data are placed in the
overlay in a deterministic manner.] {
  \includegraphics[scale=0.25]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various peer-to-peer architectures.}
\label{figure:p2p-archs}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
%
%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}

%
% TODO: SOME DISCUSSION ON THE ADVANTAGES OF UNSTRUCTURED NETWORKS FROM
%       \cite{z-yk_ddno_2005}
%
%Unstructured P2P networks o?er a number of
%important advantages: (i) An unstructured network
%imposes very small demands on individual nodes,
%and more speci?cally it allows nodes to join or leave
%the network without signi?cantly a?ecting the sys-
%tem performance. (ii) Unstructured networks are
%appropriate for content-based retrieval (e.g., key-
%word searches) as opposed to object identi?er loca-
%tion of structured overlays. (iii) Finally unstructured
%networks can easily accommodate nodes of vary-
%ing power. Consequently, they scale to very large
%sizes and they o?er more robust performance in
%the presence of node failures and connection
%unreliability.
%

@@

One of the major issues that defines the efficiency of an overlay network is the
mapping of the overlay structure to the underlying infrastructure. Recall,, in
Figure~\ref{figure:overlay} nodes $A$ and $B$ are on the same local network,
while $C$ and $D$ are in different networks. The top layer represents the
overlay interconnection formed by these nodes at a higher level. Link
connections there, can change as needed by the running environment, without any
particular constraint by the underlying physical topology. Consider the following
example. 
Assume nodes $A$, $B$, $C$ and $D$ are
connected via the physical network shown in Figure~\ref{figure:phys}, where
the network costs are given in milliseconds, and peer $A$ sends a message to
peer $D$.

\begin{figure}
\centering
  \includegraphics[scale=0.7]{img/pdf/example-physical.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}

Depending on the configuration of the overlay,
peers will experience different performance.
For example, in the overlay depicted in
Figure~\ref{figure:overlay-1}, the
message will traverse the following sequence of links in the physical layer
(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
\rightarrow 5 \rightarrow 4$. In the alternative overlay of
Figure~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow
4$. The total cost of the first sequence is $45 ms$, while the second yields a
mere $15 ms$. Therefore, we can conclude that the second overlay is more
congruent with the underlying physical network than the first one and, thus, is
more efficient.

\begin{figure}[ht]
\centering
\subfigure[Nodes B and C have direct overlay connection.] {
  \includegraphics[scale=0.5]{img/pdf/example-overlay-inefficient.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes A and D have direct overlay connection.] {
  \includegraphics[scale=0.5]{img/pdf/example-overlay-efficient.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

Ideally, an overlay network should be constructed to achieve an optimal 
mapping between the overlay paths and the underlying network links to avoid
inefficient states where redundant physical resources are utilised for the
operation of the same link in the overlay. The problem of constructing an optimal
overlay is referred to as the \emph{topology mismatch problem}, and is formally
defined as follows:
\begin{definition}
Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
problem is to construct a minimal spanning tree, where node degree is
restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols did not make use of the optimal
mapping with the underlying network topology. Gnutella, for example, is
considered far from scalable \cite{ritter_gnucantscale_2001} because every peer
chooses its neighbors in the overlay without any knowledge of the underlying
network, resulting in a mismatch between the two topologies. Queries may be
flooded over multiple paths merging in one peer while one of the paths would
suffice.  Moreover, peers may forward the same message to each other
before they receive the query messages from each other.
\cite{matei_mapgnutella_2002} showed that, even when 95\% of any two nodes are
less than 7 hops away from each other, a \emph{flooding-based} algorithm, like
the one used by Gnutella, can generate 330TB/month in a 50.000 node network. A
topology mismatch can thus have a serious effect on Internet router performance.

Similar problems are observed in decentralized structured schemes as well.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability, and robustness for the overlay. Unfortunately, this
randomness has a negative impact on the \emph{routing locality} of the network.
This means that even though the target node can be reached with a logarithmic
number of overlay hops, the distance traveled in the underlying network, during the
overlay routing process, can be far from optimal.  The ratio of the actual IP
network distance a query travels to reach an object of interest (via overlay
routing) to the minimal IP network distance to that object (i.e. through IP) is
known as \emph{stretch} or \emph{Relative Delay Penalty (RDP)} \cite{CRZ2000}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation and Goal for this Survey}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A topology unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. For example, a single logical point-to-point link on the
overlay typically corresponds to multiple physical links in the
underlying layer. Additionally, a link in the underlying network often lies
on several overlay paths causing increase of the traffic on the
link, known as the link's \emph{stress}~\cite{CRSZ2002}. Even in the case
of an approach that provides mechanisms to a new-coming node to detect close-by
peers at join time, the stochastic behavior with which peers join and leave
the overlay network, called the network's churn, further undermines the effort
for maintaining the best position in the network throughout the life-time of a
peer.

The topology mismatch problem is exacerbated by the extremely large amounts of
traffic generated by highly popular P2P applications (refer to the discussion
made in Section~\ref{section:intro}). Thus, solving the topology mismatch
problem is a fundamental challenge in contributing to the overall efficiency of
the Internet. Unfortunately, topology mismatch is known to be an NP-Hard problem
\cite{NPBOOK,C2000}. Adding to the complexity of the problem
is the fact that end-to-end latencies demonstrate \emph{triangle inequality
violations (TIVs)} which are a consequence of the Internet's structure and
routing policies and thus will remain a property of the Internet for the
foreseeable future \cite{zheng_irprtt_2005}. TIVs affect both network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
systems and makes the construction of IP-topology-aware overlays difficult.

>>>>

In this section, we provide some background information on the types of
\p\ architectures that have been implemented and deployed. We also
describe and motivate more formally the topology mismatch problem 
whose various aspects have been tackled over the past several years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (\p) Overlay Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Circa 2000, the first file-sharing \p-system introduced the \emph{servent}
concept~\cite{gnutella}, a \emph{portmanteau} that blends the notions of server
and client to denote the dual role of a node participating in \p-network. 
Such server-less systems have proven to be able to achieve outstanding
aggregate resource capacities as more and more participants join the system
without requiring additional expenditure for infrastructure. 
While certain undesirable features such as 
free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
the distribution of illegal content and other~\emph{socio-technical}
issues \cite{hughes_socp2p_2008}, have arisen, \p\ systems have
continued to gain popularity throughout the years primarily 
because they have successfully supported popular applications 
(i.e. file-sharing) amongst \emph{vast} numbers of participating users.
%%
For over a decade, excitement over the immense potential of \p\ systems has
generated a flurry of research and development, resulting in a wide range of
popular protocols, networks, and applications. As \p\ systems evolved, three
main architectures have emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

Designers of \emph{centralized} architectures 
(Figure~\ref{figure:p2p-archs:centralized}) were the first 
to recognize that requests for resources 
(e.g., CPU cycles or popular content) need not be sent to
a dedicated server. Instead, requests could be handled by 
many hosts that already possessed the resources in question. 
The downside of this approach was that it required 
centralized search that became a single point of failure, a
scalability bottleneck, and a target of malicious including \emph{DoS} attacks. 
% \emph{Denial-of-Service (DoS) attacks}).

{\sl Napster} was the most successful incarnation of the
centralized approach in file-sharing~\cite{napster}.
It maintained a \emph{central index server} based on file-lists 
that were peer-provided.
Users would query the index server that would ultimately 
yield pointers to the actual content.
Thus, by centralizing search while distributing downloads,  {\sl Napster}
achieved a highly functional design that, at its height, attracted $26$ million
users sharing approximately $80$ million songs\cite{jmm_naptopusage_2001}.
%%
The central indexing played a vital role in the demise of 
{\sl Napter} as the \emph{Recording
Industry Association of America} successfully argued in court 
copyright infringement of the artists songs traded in the system. 
%
% Napter's  centralized nature was what ultimately brought it down after Recording
% Industry Association of America (RIAA) filed a lawsuit for copyright infringement
% of the intellectual property of artists whose music was distributed by the
% Napster application. RIAA was able to point to the centralized indexing
% component of Napster as the entity responsible for illegal song-trading.

Architectures that distributed both search and resource provision 
capabilities (Figure~\ref{figure:p2p-archs:unstructred-full}) soon followed.
In these \emph{decentralized} approaches, 
resource placement within the overlay topology is random~\cite{YG-M2002};
for this reason, such architectures are referred to as \emph{unstructured}. 
Key properties  of such \p\ systems are that they
support inherent heterogeneity of peers, 
are highly resilient to peer failures,
and incur low maintenance overhead at handling 
the dynamics of peer participation \cite{stutzbach_churn_2006}. 
Unstructured systems are also known as 
\emph{broadcast-based} systems as they use message flooding among
peers to propagate search queries. 
In its first realization,
{\sl Gnutella}(\emph{v.0.4}) is a well-known example of the fully
decentralized unstructured approach.

A key disadvantage of the unstructured approach is that message flooding burdens
the network since queries travel within the network randomly, visiting nodes
that do not have the sought resource and thus wasting bandwidth.  
To save on bandwidth, unstructured networks typically limit how long 
a query travels within the overlay via the \emph{time-to-live} parameter
(\emph{TTL}). 
When a query is received by a node, it decreases its \emph{TTL} by one unit. 
If the \emph{TTL} has not reached zero, the node forwards the query 
to its neighbors.  Otherwise, it drops the query. 
While this approach prevents queries from visiting the entire network, it
does not guarantee that the resource of interest, (e.g., a rare recording) 
% of a musical group from the 1930s) 
will be ultimately found.
%%%
\begin{figure}[ht]
\centering
\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers query the centralized
entity (dotted lines) and then establish the connection to the data source.] {
  \includegraphics[scale=0.20]{img/pdf/centralized.pdf}
  \label{figure:p2p-archs:centralized}
}\qquad\qquad
\subfigure[Fully unstructured ({\sl Gnutella} \emph{v.0.4}). Data is distributed randomly,
the topology is also random and all participants are equal.] {
  \includegraphics[scale=0.20]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructred-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured ({\sl Gnutella} \emph{v.0.6} or {\sl FastTrack}). Certain
peers (bigger circles) shoulder specific responsibilities (e.g., data
indexing).] {
  \includegraphics[scale=0.20]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured ({\sl Chord} and {\sl Kademlia}). 
Peers and data are placed in the overlay in a deterministic manner.] {
  \includegraphics[scale=0.20]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various \p\ architectures.}
\label{figure:p2p-archs}
\end{figure}

To shorten search times, improve scalability, and lighten network load, 
unstructured configurations evolved into \emph{hierarchical} systems 
that dynamically assigned indexing functions to special peers, 
called \emph{ultra-}  or \emph{super-peers} 
(Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). 
Peers with reliable network connections and/or high compute power 
became \emph{ultra/super-peers}
facilitating less-powerful \emph{leaf-nodes} to find resources of
interest. 
Examples of such hierarchical unstructured systems include
{\sl Gnutella}(\emph{v.0.6})/{\sl Limewire}~\cite{gnutella}, 
{\sl KaZaA}~\cite{kazaa} and {\sl Skype}~\cite{skype}. 
Although the hierarchical approach relieves network pressure as
search queries were flooded over the much smaller subset of 
ultra/super-peers, the disadvantages of reduced search scope and 
inability to locate rare resources remained.

The \emph{decentralized structured schemes}
(Figure~\ref{figure:p2p-archs:structured}) sought to address the 
disadvantages of the above decentralized unstructured \p--approaches.
Here, the main objective was to offer 
a self-organizing infrastructure for large-scale \p\ applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
\emph{Decentralized structured schemes}
implement a \emph{Distributed Hash Table (DHT)} that maps objects 
to nodes through a deterministic mechanism. 
\emph{DHT}s provide a guaranteed bound on the
number of overlay routing hops that have to be ``traveled''
% crossed 
to locate a resource, even in the case when only a \emph{single} copy exists.  
This requires
$O$($log(n)$) hops, compared to unstructured networks that require
$O(n)$ to reliably locate any object. 
Unfortunately this efficiency does not come without cost. 
Structured overlays can only support exact-match queries 
(i.e., queries that identify resources by name) as opposed
to content-based retrieval provided by unstructured overlays. 
This means that unstructured overlays can support more versatile 
resource location queries such as keyword and partial matching searches. 
Several popular file-sharing
\p\ applications have been based on \emph{DHT}s including
{\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
{\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002},  and 
some flavors of {\sl BitTorrent}~\cite{c_bittorrent_2003}.  
Moreover, \emph{DHT}s sparked
a flurry of research for many diverse of applications ranging
from distributed search engines to event-notification and cloud-based
platforms~\cite{kbc_oceanstore_2000,rkcd_scribe_2001,mgpj_cloudsnap_2011}.

%
% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
%
%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}

%
% TODO: SOME DISCUSSION ON THE ADVANTAGES OF UNSTRUCTURED NETWORKS FROM
%       \cite{z-yk_ddno_2005}
%
%Unstructured P2P networks o?er a number of
%important advantages: (i) An unstructured network
%imposes very small demands on individual nodes,
%and more specifically?cally it allows nodes to join or leave
%the network without signi?cantly a?ecting the sys-
%tem performance. (ii) Unstructured networks are
%appropriate for content-based retrieval (e.g., key-
%word searches) as opposed to object identi?er loca-
%tion of structured overlays. (iii) Finally unstructured
%networks can easily accommodate nodes of vary-
%ing power. Consequently, they scale to very large
%sizes and they o?er more robust performance in
%the presence of node failures and connection
%unreliability.
%

@@

One of the major issues that defines the efficiency of 
an overlay network is its mapping to 
the underlying physical infrastructure. 
Recall, in Figure~\ref{figure:overlay} nodes $A$ and $B$ are 
on the same local network, while $C$ and $D$ are in different networks. 
The top layer represents the overlay interconnection formed by 
these four nodes at a higher level. 
Link connections there, can change as needed by 
the running environment, without any
particular constraint by the underlying physical topology. 
%%
For instance, assume nodes $A$, $B$, $C$ and $D$ being 
connected via the physical network shown in Figure~\ref{figure:phys}
in which costs are depicted in milliseconds. 
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/example-physical.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}
%%
When peer $A$ sends a message its $D$ counterpart
and depending on the overlay configuration, 
nodes will experience different performance.
For example, in the overlay depicted in
Figure~\ref{figure:overlay-1}, the
message will traverse the following sequence of links in the physical layer
(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
\rightarrow 5 \rightarrow 4$. In the alternative overlay of
Figure~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow
4$. 
The respective costs are $45 ms$ and $15 ms$.
% The total cost of the first sequence is $45 ms$, while the second yields a
% mere $15 ms$. 
Evidently, the second overlay is more
congruent with the underlying physical network in comparison 
to the first and, thus, is more efficient.

\begin{figure}[ht]
\centering
\subfigure[Nodes $B$ and $C$ have direct overlay connection.] {
  \includegraphics[scale=0.4]{img/pdf/example-overlay-inefficient.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes $A$ and $D$ have direct overlay connection.] {
  \includegraphics[scale=0.4]{img/pdf/example-overlay-efficient.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

Ideally, an overlay should be constructed to achieve an \emph{optimal} 
mapping between the overlay paths and the underlying network links to avoid
inefficient states where redundant physical resources are utilized for the
operation of the same link in the overlay. 
The problem of constructing an optimal
overlay is referred to as the \emph{topology mismatch problem}, and is formally
defined as follows:
\begin{definition}
Let $V$=$\{v_1$, $\dots$, $v_n\}$ be a set of points denoting the network nodes,
$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
problem is to construct a minimal spanning tree, where node degree is
restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols did not make use of the optimal
mapping with the underlying network topology. {\sl Gnutella}, for example, is
considered far from scalable~\cite{ritter_gnucantscale_2001} since every peer
chooses its neighbors in the overlay without any knowledge of the underlying
network, resulting in mismatch. 
Queries may be flooded over multiple paths merging in one peer 
while one of the paths would suffice.  
Moreover, peers may forward the same message to each other
before they receive the query messages from each other.
\cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
are less than $7$ hops away from each other, 
a \emph{flooding-based} {\sl Gnutella}-like algorithm 
can generate $330$TB/month in a $50,000$ node network. 
A topology mismatch can thus have a serious effect 
on Internet router performance.

Similar problems are observed in decentralized structured schemes as well.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability, and robustness for the overlay. Unfortunately, this
randomness has a negative impact on the \emph{routing locality} of the network.
This means that even though the target node can be reached with a logarithmic
number of overlay hops, the distance traveled in the underlying network, 
during the overlay routing process, can be far from optimal.  
%%AD the above text is disconnected with what follows just below... 
%%	- does it give a metric for the "optimality" we seek???
The ratio of the actual \emph{IP} network distance a query travels to 
reach an object of interest (via overlay routing) to 
the minimal \emph{IP} network distance to that object 
(i.e. through \emph{IP}) is known as \emph{stretch} 
or \emph{Relative Delay Penalty (RDP)}~\cite{CRZ2000}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation and Goal for this Survey}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A topology--unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. 
In particular, a single logical point-to-point link on the
overlay typically corresponds to multiple physical links in the
underlying layer. 
Moreover, a link in the physical network often lies
on several overlay paths inadvertently increasing the traffic on the
link; this is known as the link's \emph{stress}~\cite{CRSZ2002}. 
There is also the stochastic phenomenon during which peers 
depart and others arrive in the network, known 
as \emph{churn}~\cite{stutzbach_churn_2006}.
Even when an approach offers the best possible location
when a node joins in, the churn is set to  
continuously undermine the effort to maintain the best possible position
of the  peer in the network.
%%AD the paragraph below is ... LONG - rephrased as above...
% Even in the case of an approach that provides mechanisms 
% to a new-coming node to detect close-by peers at join time, 
% the stochastic behavior with which peers join and leave
% the overlay network, called the network's churn, 
% further undermines the effort for maintaining 
% the best position in the network throughout the life-time of a peer.
